---
title: "Proteomic data processing"
author: "Camilo Posso"
date: "5/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(tibble)
```

## Data formats

Using the data available in the crosstabs and a second metadata file, we can create a Synapse table that is in 'long form' and has all the protein/phosphosite values together with their relevant metadata.

### Issues with SSL? Try disconnecting your VPN

```{r synapselogin}
source("../../util/synapseUtil.R")

syn <- synapseLogin()

tables.syn <- c("Phospho peptide corrected" = "syn25681780",
                "Phospho peptide uncorrected" = "syn25681781",
                "Phospho peptide uncorrected phospho coefficients" = "syn25681782",
                "Phospho siteID corrected" = "syn25681783",
                "Phospho siteID uncorrected" = "syn25681786",
                "Phospho siteID uncorrected phospho coefficients" = "syn25681787")

metadata.syn <- "syn25681625"

```


## Merge metadata and Phospho proteomics data


```{r pressure, echo=FALSE}
meta.tab <- readxl::read_xlsx(syn$get(metadata.syn)$path)
meta.tab$`PNNL ID` <- sub("^PTRC_Ex17_(.*)$","\\1",meta.tab$`PNNL ID`)
index <- which(names(meta.tab) == "Tx Status")
names(meta.tab)[index] <- "Description"

print(head(meta.tab))

tables <- list()

for (i in 1:length(tables.syn)){
  name <- names(tables.syn)[[i]]
  crosstab.syn <- tables.syn[[i]]
  mat <- read.csv(syn$get(crosstab.syn)$path,sep='\t',header=T)
  names(mat) <- sub("^X(.*)$","\\1",names(mat))
  mat <- add_column(mat, Accession = rownames(mat), .before = 1)
  

  mat <- mat%>%tidyr::pivot_longer(cols=c(2:ncol(mat)),names_to='Sample',
                      values_to='LogRatio')
  mat <- left_join(mat,meta.tab, by = c("Sample" ="PNNL ID")) %>% subset(!is.na(LogRatio))
  tables[[name]] <- mat
}
```

Now we have all the data in long form.  The last thing that needs to be done is to store them both to a synapse table. This requires knowing the synapse project.

```{r store tables}
for (name in names(tables.syn)){
  print(name)
  mat <- tables[[name]]
  name <- paste("PTRC_ex17",name,sep = " ")
  print(name)
  synTableStore(mat,name,parentId='syn22128879')
}
```

